---
title: "Stats_challenge_repro_30_05"
author: "Doris"
date: "5/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
library(ggplot2)
library(ggpubr)
library(dplyr)
library(tidyr)
```


```{r start}
data_cats = read.csv("~/Propulsion/Statistics/DataSets/Cat_data.csv",header = TRUE)
str(data_cats)
```

### Data exploration

- show 
```{r pressure, echo=FALSE}
number_breed = count(data_cats, Breed_group)
number_weaning = count(data_cats, Weaning_age)

ggplot(number_breed, aes(x=Breed_group, y = n))+
  geom_bar(stat = 'identity', color = 'grey')+
  ggtitle('Number of cats per breed')+
  xlab('Breed groups')+
  ylab('Number of cats')+
  theme_bw()


# add labels!
ggplot(number_weaning, aes(x=Weaning_age, y = n))+
  geom_bar(stat = 'identity', fill = c('red','red','red','grey','grey','grey','green','green'))+
  scale_x_continuous(breaks = c(1,2,3,4,5,6,7,8),
                     labels = c('<8 weeks','8-9 weeks','10-11 weeks','12-13 weeks',
                                '14-15 weeks','16 w - 1 year','adult','never'))+
  #ggtitle('Number of cats per weaning age')+
  xlab('Weaning age')+
  ylab('Number of cats')+
  theme_bw()



ggplot(data_cats, aes(x=Age)) +
	geom_histogram(binwidth = 1, col = "grey")+
  xlab('Cat age [years]')+
  ylab('Number of cats')+
  ggtitle('Age distribution of analyzed cats')+
  theme_bw()



```

### Chi squared test between 9 environmental factors and 'Agression towards family members'


```{r}
data_cats$aggression_binary = rep(0,count(data_cats))
data_cats$aggression_binary[data_cats$Aggression_owner > 1] = 1

mean(data_cats$Age)
median(data_cats$Age)
data_cats$age_binary = rep(0,count(data_cats))
data_cats$age_binary[data_cats$Age > 3.88] = 1

# plot aggresion beh 1-5, and in same plot the distr of binary


```

```{r}
# Null hypothesis: there is no relation between weaning and aggression
# contingency table
table_WA = table(data_cats$Weaning_age, data_cats$aggression_binary)
test_WA = chisq.test(table_WA, correct = TRUE)
row_WA = c(test_WA$statistic[1], test_WA$p.value, test_WA$parameter[1])

# Null hypothesis: there is no relation between sex and aggression
table_sex = table(data_cats$Gender, data_cats$aggression_binary)
test_sex = chisq.test(table_sex, correct = TRUE)
row_sex = c(test_sex$statistic[1], test_sex$p.value, test_sex$parameter[1])

# Null hypothesis: there is no relation between age and aggression
table_age = table(data_cats$Age, data_cats$aggression_binary)
test_age = chisq.test(table_age, correct = TRUE)
row_age = c(test_age$statistic[1], test_age$p.value, test_age$parameter[1])

table_horm = table(data_cats$Neuter_status, data_cats$aggression_binary)
test_horm = chisq.test(table_horm, correct = TRUE)
row_horm = c(test_horm$statistic[1], test_horm$p.value, test_horm$parameter[1])

table_breed = table(data_cats$Breed_group, data_cats$aggression_binary)
test_breed = chisq.test(table_breed, correct = TRUE)
row_breed = c(test_breed$statistic[1], test_breed$p.value, test_breed$parameter[1])

table_out = table(data_cats$Outdoors, data_cats$aggression_binary)
test_out = chisq.test(table_out, correct = TRUE)
row_out = c(test_out$statistic[1], test_out$p.value, test_out$parameter[1])

table_other = table(data_cats$Other_cats, data_cats$aggression_binary)
test_other = chisq.test(table_other, correct = TRUE)
row_other = c(test_other$statistic[1], test_other$p.value, test_other$parameter[1])

table_shy = table(data_cats$Shyness_component, data_cats$aggression_binary)
test_shy = chisq.test(table_shy, correct = TRUE)
row_shy = c(test_shy$statistic[1], test_shy$p.value, test_shy$parameter[1])

our_chi_results = data.frame('Weaning_age'=row_WA, 'Sex'=row_sex,'Age'=row_age, 'Hormonal_status'=row_horm,'Breed'=row_breed,'Access_to_outdoors'=row_out, 'Other_cats'=row_other,'Shyness'=row_shy)


our_chi_results = t(our_chi_results)
chi_results_df = as.data.frame(our_chi_results)
names(chi_results_df)[names(chi_results_df) == 'V2'] = 'p_value'
chi_results_df$`X-squared` = round(chi_results_df$`X-squared`, digits = 2)

chi_results_df
```

### Other tests for continuous data --> binary output

```{r age}
shapiro.test(sample(data_cats$Age, 5000, replace=FALSE))
hist(data_cats$Age)
# not normal dist!

wilcox.test( data_cats$Age ~ data_cats$aggression_binary, data_cats, paired = FALSE)
wilcox.test( data_cats$Age ~ data_cats$aggression_binary, data_cats)

# violin plot
violinplot_aggr = ggplot(data_cats, aes(x = data_cats$aggression_binary, y = data_cats$Age, group=data_cats$aggression_binary))+
	geom_violin(draw_quantiles = TRUE)+
  scale_x_continuous(breaks = c(0.0,1.0), labels = c('No aggressions','Aggressions'))+
  ylab('Cat age [years]')+
  xlab(' ')+
	theme_bw()

# better than violin plots
better_violinplot_aggr = ggplot(data_cats, aes(x=data_cats$aggression_binary, y=data_cats$Age,group=data_cats$aggression_binary)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 0.2, dotsize=0.3, position=position_dodge(0.8), alpha=0.2)+
  scale_x_continuous(breaks = c(0.0,1.0), labels = c('No aggressions','Aggressions'))+
  #ggtitle('Age distributions of normal vs. aggressive cats')+
  ylab('Cat age [years]')+
  xlab(' ')+
	theme_bw()

theme_set(theme_pubr())

violin_fig = ggarrange(violinplot_aggr, better_violinplot_aggr,
                    labels = c("Violin plot", 'Dot plot'),
                    ncol = 2, nrow = 1)
violin_fig


mean_diff_real = mean(data_cats$Age[data_cats$aggression_binary == 0]) - mean(data_cats$Age[data_cats$aggression_binary == 1])

median(data_cats$Age[data_cats$aggression_binary == 0])
median(data_cats$Age[data_cats$aggression_binary == 1])

# try permutation test

data_permut = data_cats[ ,c('Age',"aggression_binary")]
     
list_mean_diff = vector()
# loop 1000 times
for (i in 1:1000){ 
  data_permut$aggression_binary = sample(data_permut$aggression_binary, length(data_permut$aggression_binary), replace = FALSE)
  curr_mean_diff = mean(data_permut$Age[data_permut$aggression_binary == 0]) - mean(data_permut$Age[data_permut$aggression_binary == 1])

  list_mean_diff = append(list_mean_diff,curr_mean_diff)
  
  }

mean_perm = mean(list_mean_diff)


# make df for ggplot!
my_df = data.frame('dist_mean' = list_mean_diff)
  
ggplot(my_df)+
  geom_histogram(mapping = aes(x=dist_mean), binwidth = 0.1)+
  geom_vline(xintercept=mean_diff_real, color="green")+
  geom_vline(xintercept=mean_perm, color="red")+
  ggtitle('Sample distribution of mean weights')+
  theme_bw()


```

```{r shyness}
shapiro.test(sample(data_cats$Shyness_component, 5000, replace=FALSE))
hist(data_cats$Shyness_component)

wilcox.test( data_cats$Shyness_component ~ data_cats$aggression_binary, data_cats, paired = FALSE)

# ggplot(data_cats, aes(x = data_cats$aggression_binary, y = data_cats$Shyness_component, group=data_cats$aggression_binary))+
# 	geom_dotplot(aes(y = data_cats$Shyness_component))+
#   xlab('No aggression vs aggression')+
#   ylab('Cat shyness component')+
# 	theme_bw()

# better than violin plots
ggplot(data_cats, aes(x=data_cats$aggression_binary, y=data_cats$Shyness_component, group=data_cats$aggression_binary)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth = 0.05, dotsize=0.3, position=position_dodge(0.8), alpha=0.2)+
  scale_x_continuous(breaks = c(0.0,1.0), labels = c('No aggressions','Aggressions'))+
  ggtitle('Shyness component distributions of normal vs. aggressive cats')+
  ylab('Cat shyness component')+
  xlab(' ')+
	theme_bw()

```

### comparison between weaning age groups, which ones are signif more aggressive?

```{r}
# subsample 100 and calculate mean
group_1 = data_cats[(data_cats$Weaning_age == 1 | data_cats$Weaning_age == 4), c(6,22)]
group_2 = data_cats[(data_cats$Weaning_age == 2 | data_cats$Weaning_age == 4), c(6,22) ]
group_3 = data_cats[(data_cats$Weaning_age == 3 | data_cats$Weaning_age == 4), c(6,22) ]
group_4 = data_cats[(data_cats$Weaning_age == 5 | data_cats$Weaning_age == 4), c(6,22) ]

group_5 = data_cats[(data_cats$Weaning_age == 5 | data_cats$Weaning_age == 1 | data_cats$Weaning_age == 2 | data_cats$Weaning_age == 3), c(6,22)] 
# group 1-3, call early_weaned
group_5$Weaning_age[group_5$Weaning_age < 4] = 'early_weaned'

group_6 = data_cats[ , c(6,22)]
# group 7 and 8 as not weaned, rest as weaned
group_6$Weaning_age[group_6$Weaning_age > 6] = 'not_weaned'
group_6$Weaning_age[group_6$Weaning_age < 7] = 'weaned'

test_group_1 = chisq.test(table(group_1$Weaning_age, group_1$aggression_binary), correct = TRUE)
test_group_2 = chisq.test(table(group_2$Weaning_age, group_2$aggression_binary), correct = TRUE)
test_group_3 = chisq.test(table(group_3$Weaning_age, group_3$aggression_binary), correct = TRUE)
test_group_4 = chisq.test(table(group_4$Weaning_age, group_4$aggression_binary), correct = TRUE)
test_group_5 = chisq.test(table(group_5$Weaning_age, group_5$aggression_binary), correct = TRUE)
test_group_6 = chisq.test(table(group_6$Weaning_age, group_6$aggression_binary), correct = TRUE)

col1 = c(test_group_1$statistic[1], test_group_1$p.value, test_group_1$parameter[1])
col2 = c(test_group_2$statistic[1], test_group_2$p.value, test_group_2$parameter[1])
col3 = c(test_group_3$statistic[1], test_group_3$p.value, test_group_3$parameter[1])
col4 = c(test_group_4$statistic[1], test_group_4$p.value, test_group_4$parameter[1])
col5 = c(test_group_5$statistic[1], test_group_5$p.value, test_group_5$parameter[1])
col6 = c(test_group_6$statistic[1], test_group_6$p.value, test_group_6$parameter[1])

weaning_results = data.frame('Wean 1 vs 4'=col1, 'Wean 2 vs 4'=col2, 'Wean 3 vs 4'=col3, 'Wean 5 vs 4'=col4, 
                             'Wean 5 vs early'=col5, 'Not weaned vs weaned'=col6)
                             
weaning_results = as.data.frame(t(weaning_results))

names(weaning_results)[names(weaning_results) == 'V2'] = 'p_value'
weaning_results$`X-squared` = round(weaning_results$`X-squared`, digits = 2)
weaning_results = as.data.frame(t(weaning_results))

weaning_results[3,] = 1
weaning_results

```

#### Power analysis

n = 5726

#### regression model:

start model: aggression as function of weaning age, age, sex

final model: aggression as function of weaning age, age, sex, breed

```{r}
start_model = glm(data_cats$aggression_binary ~ data_cats$Weaning_age + data_cats$Age, family = binomial(link='logit'))

summary(start_model)
```

```{r}
breed_model = glm(data_cats$aggression_binary ~ data_cats$Gender + data_cats$Age + data_cats$Weaning_age + data_cats$Breed_group,family = binomial(link='logit'))

summary(breed_model)
```

```{r}
test_model = glm(data_cats$aggression_binary ~ data_cats$Age + data_cats$Gender, family = binomial(link='logit'))

summary(test_model)

```


